\emph{Cree}: A modern toolbox for readymade economic experiments

### 2015-06-22 Latex/paper.tex
Fixed typos

\Cree is build using web technologies. Content is structured using the markup language \emph{HTML}. The layout and presentation across different screen sizes is archived with the style sheet language \emph{CSS}. And all logic is constructed using the programming language \emph{JavaScript}. All web browsers interpret these three cornerstone languages and render pages accordingly. Subjects participate in \Cree experiments though a web browser. Because \Cree is fundamentally native to the web browser, it avoids many of the restrictions that other toolboxes suffer from.

@@

Technological advances opens up a new path, however the path is not without obstacles. This paper explores and discusses how to handle these obstacles. Some obstacles are alleviated through appropriate technical design of the toolbox. Other obstacles require actions taken by the researcher.

@@

Subjects access \Cree experiments though a web browser. Any modern web browser will suffice\footnotemark[1], which means that subjects can participate in experiments using smart-phones, tablets, or any computer (Windows, Mac, Linux). Every subject will need a device. The researcher can choose to provide devices, or have subjects use their own personal devices.

@@

The second category contains toolboxes written in \emph{Python}. Python is platform independent, so these toolboxes can run on several operating systems, but they require that Python be installed on the equipment facing subjects\footnotemark[2]. These toolboxes also require downloading software.

@@

It is commonly assumed that subjects are \emph{naïve}, eg. encounter the experimental material for the first time, and have no a priori knowledge of the experiment, the purpose, nor knowledge of related experiments. This is not true for experiments conducted over the internet, and is often misapprehended by researchers, as shown by \citet*{Chandler_Mueller_Paolacci_2014}. They investigate experiments where subjects are recruited through Amazon's \emph{Mechanical Turk} system. \emph{Mechanical Turk} is a crowdsourcing platform of 500,000 workers that perform mostly small tasks and are paid a piece rate. Every worker has an account, and \emph{Mechanical Turk} prevents an account from performing the same task twice. \cite{Chandler_Mueller_Paolacci_2014} find that:

\begin{enumerate}
\item subjects have experience from related experiments. By pooling 16,408 observations (or completed tasks) across 132 studies, they find just 7,498 unique subjects -- averaging 2.24 observation per subject. The 10\% most prolific subjects account for 41\% of all observations. In addition \cite{Chandler_Mueller_Paolacci_2014} survey 300 workers on \emph{Mechanical Turk} and find that 56\% have previously been exposed to the \emph{Prisoner's dilemma} and 52\% to the \emph{Ultimatum game}. 
\item subjects share information about experiments. About one fourth of workers in the \cite{Chandler_Mueller_Paolacci_2014} survey respectively report they know other workers in person, and that they read or discuss \emph{Mechanical Turk} online on blogs and forums. The focus of the conversation is typically not the specific contents of the experiment, but rather pay rates and reputation. \cite{Chandler_Mueller_Paolacci_2014} conclude that researchers should be particularly aware of cross talk, when information can increase financial rewards -- such as when using techniques to increase attention (\emph{instructional manipulation checks}, or \emph{got-cha} questions), or when using techniques to deny payment. They suggest that researchers ask subjects how they found the experiment, and monitor discussion boards that refer a lot of subjects to the experiment.
\item few subjects attempt to participate repeatedly in the same experiment. Amazon forbids multiply or \emph{sock-puppet} accounts, and actively enforces this policy. \cite{Chandler_Mueller_Paolacci_2014} find 1.2\% of subjects share IP addresses as well as demographic characteristics -- eg. possible \emph{sock-puppet} accounts. They attribute the low percentage to the active enforcement by Amazon.
\end{enumerate}

@@

\subsection{Multiple versions}
% How can Cree run on smart-phones, tablets, and computers all at the same time?

@@

There are some misconceptions that I would like to address. 
\begin{enumerate}
\item The main purpose of a toolbox is not to offer comfort or convenience to researchers, but to offer the necessary tools. Convenience is secondary.
\item There is a trade-off between usability and flexibility. The programmer bases every improvement in usability on a judgement call. Usability is improved when the set of possible choices are reduced, this in turn reduces flexibility. Thus a toolbox should be evaluated on its ability to balance the trade-off between usability and flexibility. 
\item All toolboxes essentially require some programming skills. It is true that some toolboxes allow researchers to run preprogrammed experiments, without any additional programming. But once the researcher starts building own experiments, or significantly modifies existing experiments, programming is unavoidable. A programming language is nothing more than human logic and instructions in a machine interpretable format, and this is why programming is unavoidable when researchers build logic into their experiments.
\end{enumerate}

@@

This paper introduces \Cree, a modern toolbox to design and conduct economic experiments. \Cree places very few restrictions on the devices facing subjects, thus relinquishing economic experiments from the grips of the computer-equipped laboratory. In turn economic experiments designed in \Cree can also run in \emph{extra-laboratory} settings such as classrooms, workplaces and over the internet. While \Cree provides researchers with new tools and opportunities, it also opens up for new issues and obstacles. This paper shows how many of these obstacles are solve in the technical design of \Cree, while a few obstacles remain that require action by the researcher. The main remaining issue facing researchers when transferring from a laboratory setting to an \emph{extra-laboratory} setting is not subject attention, subject compensation, nor the quality of the results. It is recruitment and specifically the selection process. If the experiment is designed appropriately the first set of issues resolve themselves. However the method used to recruit subjects, especially if subjects self-select into the experiment, may change the mindset, behaviour, validity and ultimately the results of the experiment. The selection process requires a third-party prescreening system to manage the pool of subjects. And it may require running prescreen surveys to appropriately exclude subjects on a priori grounds. The problems surrounding the selection process are not exclusive to internet experiments, just more prevalent in this setting.

>>>>

\Cree is build using web technologies. Content is structured using the markup language \emph{HTML}. The layout and presentation across different screen sizes is archived with the style sheet language \emph{CSS}. And all logic is constructed using the programming language \emph{JavaScript}. All web browsers interpret these three cornerstone languages and render pages accordingly. Subjects participate in \Cree experiments through a web browser. Because \Cree is fundamentally native to the web browser, it avoids many of the restrictions that other toolboxes suffer from.

@@

Technological advances opens up a new path, however the path is not without obstacles. This paper explores and discusses how to handle these obstacles. Some obstacles are alleviated through appropriate technical design of the toolbox. Other obstacles requires actions taken by the researcher.

@@

Subjects access \Cree experiments through a web browser. Any modern web browser will suffice\footnotemark[1], which means that subjects can participate in experiments using smart-phones, tablets, or any computer (Windows, Mac, Linux). Every subject will need a device. The researcher can choose to provide devices, or have subjects use their own personal devices.

@@

The second category contains toolboxes written in \emph{Python}. Python is platform independent, so these toolboxes can run on several operating systems, but they require that Python is installed on the equipment facing subjects\footnotemark[2]. These toolboxes also require downloading software.

@@

It is commonly assumed that subjects are naïve, eg. encounter the experimental material for the first time, and have no a priori knowledge of the experiment, the purpose, nor knowledge of related experiments. This is not true for experiments conducted over the internet, and is often misapprehended by researchers, as shown by \citet*{Chandler_Mueller_Paolacci_2014}. They investigate experiments where subjects are recruited through Amazon's \emph{Mechanical Turk} system. \emph{Mechanical Turk} is a crowdsourcing platform of 500,000 workers that perform mostly small tasks and are paid a piece rate. Every worker has an account, and \emph{Mechanical Turk} prevents an account from performing the same task twice. \cite{Chandler_Mueller_Paolacci_2014} find that:

\begin{enumerate}
\item subjects have experience from related experiments. By pooling 16,408 observations (or completed tasks) across 132 studies, they find just 7,498 unique subjects -- averaging 2.24 observation per subject. The 10\% most prolific subjects account for 41\% of all observations. In addition \cite{Chandler_Mueller_Paolacci_2014} survey 300 workers on \emph{Mechanical Turk} and find that 56\% have previously been exposed to the \emph{Prisoner's dilemma} and 52\% to the \emph{Ultimatum game}. 
\item subjects share information about experiments. About one fourth of workers in the \cite{Chandler_Mueller_Paolacci_2014} survey respectively report that they know other workers in person, and that they read or discuss \emph{Mechanical Turk} online on blogs and forums. The focus of the conversation is typically not the specific contents of the experiment, but rather pay rates and reputation. \cite{Chandler_Mueller_Paolacci_2014} conclude that researchers should be particularly aware of cross talk, when information can increase financial rewards -- such as when using techniques to increase attention (\emph{instructional manipulation checks}, or \emph{got-cha} questions), or when using techniques to deny payment. They suggest that researchers ask subjects how they found the experiment, and monitor discussion boards that refer a lot of subjects to the experiment.
\item few subjects attempt to participate repeatedly in the same experiment. Amazon forbids multiply or \emph{sock-puppet} accounts, and actively enforces this policy. \cite{Chandler_Mueller_Paolacci_2014} find 1.2\% of subjects share IP addresses as well as demographic characteristics -- eg. possible \emph{sock-puppet} accounts. They attribute the low percentage to the active enforcement by Amazon.
\end{enumerate}

@@

\subsection{Single versions}
% How can Cree run on smart-phones, tablets, and computers all at the same time?

@@

There are some misconceptions that I would like to address. 
\begin{enumerate}
\item The main purpose of a toolbox is not to offer comfort or convenience to researchers, but to offer the necessary tools. Convenience is secondary.
\item There is a trade-off between usability and flexibility. The programmer bases every improvement in usability on a judgement call. Usability is improved when the set of possible choices are reduced, this in turn reduces flexibility. Thus a toolbox should be evaluated on its ability to balance the trade-off between usability and flexibility. 
\item All toolboxes essentially requires some programming skills. It is true that some toolboxes allow researchers to run preprogrammed experiments, without any additional programming. But once the researcher starts building own experiments, or significantly modifies existing experiments, programming is unavoidable. A programming language is nothing more than human logic and instructions in a machine interpretable format, and this is why programming is unavoidable when researchers build logic into their experiments.
\end{enumerate}

@@

This paper introduces \Cree, a modern toolbox to design and conduct economic experiments. \Cree places very few restrictions on the devices facing subjects, thus relinquishing economic experiments from the grips of the computer-equipped laboratory. In turn economic experiments designed in \Cree can also run in \emph{extra-laboratory} settings such as classrooms, workplaces and over the internet. While \Cree provides researchers with new tools and opportunities, it also opens up for new issues and obstacles. This paper shows how many of these obstacles are solved in the technical design of \Cree, while a few obstacles remain that require action by the researcher. The main remaining issue facing researchers when transferring from a laboratory setting to an \emph{extra-laboratory} setting is not subject attention, subject compensation, nor the quality of the results. It is recruitment and specifically the selection process. If the experiment is designed appropriately the first set of issues resolve themselves. However the method used to recruit subjects, especially if subjects self-select into the experiment, may change the mindset, behaviour, validity and ultimately the results of the experiment. The selection process requires a third-party prescreening system to manage the pool of subjects. And it may require running prescreen surveys to appropriately exclude subjects on a priori grounds. The problems surrounding the selection process are not exclusive to internet experiments, just more prevalent in this setting.
